---
title: 'Statistical report: Use of care'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 8
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.2. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart", scale_down = FALSE)
```

For cases with 4-5 available controls 3 controls are sampled at random without replacement. 

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

The median age (IQR) is `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

```{r metanpr}
mykable(metaout, caption = "Additional outcomes from NPR")
```

```{r metalos}
mykable(metalos, caption = "Lenght of stay variables from NPR")
```

```{r metaddr}
mykable(metalm, caption = "Additional outcomes from DDR")
```

Emergency visits are defined as MVO = 046 or 100 and PVARD = 2 (either in out-patient or in-patient care). 

Length of stay is defined as total time (for the respective diagnosis in HDIA position) 
in in-patient care from index to end of follow-up. 

\clearpage
\newpage

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

Since the cases and controls are loosely matched dependency will not be incorporated into the statistical analyses. 

## Baseline characteristics

```{r, child = "../src/tab1_hfc.Rmd"}

```

```{r, child = "../src/tab1_ef.Rmd"}

```

```{r, child = "../src/tab1_efc.Rmd"}

```

\clearpage
\newpage

## Outcomes

Data were censored at 2019-12-31 or death/emigration, or, for the controls, diagnosis of HF. 

Time to first events were presented with Kaplan-Meier curves for all-cause mortality 
and cumulative incidence curves treating death as a competing event for other outcomes. 
The mean cumulative function (MCF) was used to depict the recurrent events. 

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

Cox proportional hazards regressions were 
used to model the time to first event. Recurrent events were model using a negative binomial generalized linear model 
including the log of time as an offset in the model. In the presentation by EF vs HF-free controls 3 separate models where conducted, one within each EF group.  

The median (q1-q3) follow-up is 
`r pdata %>% summarise(med = dF(median(sos_outtime_death / 365.25), dig = 1),
                                             q1 = dF(quantile(sos_outtime_death / 365.25, 0.25), dig = 1),
                                             q3 = dF(quantile(sos_outtime_death / 365.25, 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r pdata %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km_hfc.Rmd"}

```

```{r, child = "../src/km_ef.Rmd"}

```

```{r, child = "../src/km_efc.Rmd"}

```

```{r, child = "../src/out_hfc.Rmd"}

```

```{r, child = "../src/out_ef.Rmd"}

```

```{r, child = "../src/out_efischemic.Rmd"}

```

```{r, child = "../src/out_efc.Rmd"}

```

```{r, child = "../src/los.Rmd"}

```

\clearpage
\newpage

## Treatments

Treatments over time were presented graphically by EF. The crude RR was estimated using a 
Generalized Linear Model with a binomial distribution and a log link [@gee]. 

```{r, child = "../src/treat_ef.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/useofcare. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
